---
title: "promor_supplementary_figures"
output: html_document
date: "2023-01-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# promor: a comprehensive R package for label-free proteomics data analysis and predictive modeling

## Supplementary Figures


```{r upload}
library(promor)
raw1 <- create_df(
  prot_groups = "https://raw.githubusercontent.com/caranathunge/promor_example_data/main/pg1.txt",
  exp_design = "https://raw.githubusercontent.com/caranathunge/promor_example_data/main/ed1.txt"
)
raw_filtered <- filterbygroup_na(raw1, set_na = 0.4)
```

### Figure S1

Fig. S 1: A heatmap showing the missing data distribution in protein intensity data. The data consists
of protein intensity values for 4584 protein groups (rows) in six samples. The proteins on the y-axis have
been reordered by mean intensity and the samples on the x-axis by the sum of intensity to show missing
data distribution patterns in the data. Proteins with lower mean intensity appear to have higher levels of
missing data and sample H 3 have the most number of missing values. Input data from Cox et al. (2014)
downloaded from the PRIDE Archive (PRIDE ID: PXD000279).

```{r heatmapna}
heatmap_na(raw_filtered, palette = "magma", 
           reorder_y = TRUE, y_fun = mean,
           reorder_x = TRUE, x_fun =sum,
           plot_width = 3, plot_height = 6)
```


### Figure S2

Fig. S 2: Density plots showing the impact of missing data imputation on the protein intensity data
distribution in each sample. ‘minProb’ imputation method was used. Input data from Cox et al. (2014)
downloaded from the PRIDE Archive (PRIDE ID: PXD000279).

```{r imputeplot}
imp_df_mp <- impute_na(raw_filtered, method = "minProb", seed = 327)
impute_plot(original = raw_filtered, 
            imputed = imp_df_mp, 
            global = FALSE, 
            n_row = 3, n_col = 3, palette = "magma",
            plot_width = 6, plot_height = 3.5)
```


### Figure S3

**For visualization purposes only. This data set does not need to be normalized as it has been already normalized with the MaxLFQ algorithm**

Fig. S 3: Density plots showing the impact of data normalization on the protein intensity data distribution.
‘quantile’ normalization method was applied here for visualization purposes only. Input data from Cox et al.
(2014) downloaded from the PRIDE Archive (PRIDE ID: PXD000279).

```{r normeplot}
norm_df <- normalize_data(imp_df_mp)
norm_plot(original = imp_df_mp, normalized = norm_df, type = "density", palette = "magma",
          plot_width = 5, plot_height = 3.5)
     
```    



### Figure S4

Fig. S 4: Volcano plot showing the results from the differential expression analysis. Significance criteria
used: log-2-fold change of 1 and adjusted p-value of 0.05. Protein groups meeting the significance criteria
are colored in orange. Input data from Cox et al. (2014) downloaded from the PRIDE Archive (PRIDE ID:
PXD000279).


```{r volcanoplot}
fit_df <- find_dep(imp_df_mp)
volcano_plot(fit_df,
             text_size = 6,
             palette = "magma")
```   



### Figure S5

Fig. S 5: Heatmap showing the results from the differential expression analysis. The top 20 protein groups
that meet the significance criteria (log-2-fold change of 1 and adjusted p-value of 0.05) are shown here. Input
data from Cox et al. (2014) downloaded from the PRIDE Archive (PRIDE ID: PXD000279).
4


```{r heatmapde}
heatmap_de(fit_df,
           imp_df_mp,
           palette = "magma")
   
```   

### Figure S6

Fig. S 6: Scatter plots showing the correlation of intensities between pairs of technical replicates. The data
consists of protein intensity values for 839 protein groups in six samples. Input data from Ramond et al.
(2015) downloaded from the PRIDE Archive (PRIDE ID: PXD001584).

```{r corrplot}
raw2 <- create_df(
  prot_groups = "https://raw.githubusercontent.com/caranathunge/promor_example_data/main/pg2.txt",
  exp_design = "https://raw.githubusercontent.com/caranathunge/promor_example_data/main/ed2.txt",
  tech_reps = TRUE
)

corr_plot(raw2, rep_1 = 1, 
          rep_2 = 2,
          n_row = 3, 
          n_col = 2, 
          text_size = 12, 
          palette = "magma")
```         
   
### Figure S7

Volcano plot showing the results from the differential expression analysis. Significance criteria
used: log-2-fold change of 1 and adjusted p-value of 0.05. Protein groups meeting the significance criteria
are colored in orange. Input data from Ramond et al. (2015) downloaded from the PRIDE Archive (PRIDE
ID: PXD001584).

```{r volcanoplot2}
raw_ave <- aver_techreps(raw2)
raw_filtered <- filterbygroup_na(raw_ave, set_na = 0.40)
imp_df_mp2 <- impute_na(raw_filtered)
fit_df2 <- find_dep(imp_df_mp2)
volcano_plot(fit_df, 
             palette = "magma")

```
    
### Figure S8

Fig. S 8: The variation of intensities in six proteins between COVID severe and non-severe patient groups.
Input data from Suvarna et al. (2021) downloaded from the PRIDE Archive (PRIDE ID: PXD022296).
         
```{r preprocess}
covid_model_df <- pre_process(
  covid_fit_df, 
  covid_norm_df
)
feature_plot(
    model_df = covid_model_df,
    type = "density",
    n_row = 4,
    n_col = 2,
    palette = "magma"
)
```

### Figure S9

Fig S9: Dot plots showing model performance. Models built using four machine-learning algorithms are
assessed for performance using Accuracy and Kappa. Input data from Suvarna et al. (2021) downloaded
from the PRIDE Archive (PRIDE ID: PXD022296).

```{r perforplot}
covid_split_df <- split_data(model_df = covid_model_df, seed = 8314)
covid_model_list <- train_models(split_df = covid_split_df, 
                                 algorithm_list= c("svmLinear",
                                                  "naive_bayes",
                                                  "knn",
                                                  "rf"),
                                 seed = 351)
performance_plot(
  model_list = covid_model_list,
  type = "dot",
  palette = "magma"
)
```


### Figure S10

Fig S10: Lollipop plots showing variable (protein) importance. Proteins are sorted according to their
importance in the models built using four different machine learning algorithms. Input data from Suvarna
et al. (2021) downloaded from the PRIDE Archive (PRIDE ID: PXD022296).

```{r varimp}
varimp_plot(
  model_list = covid_model_list,
  text_size = 7,
  n_row = 2,
  n_col = 2,
  palette = "magma"
)
```
     
     
### Figure S11

Fig. S 11: Receiver operating characteristic curves illustrating the diagnostic ability of the models built
using four different machine learning algorithms. Area under the curve (AUC) estimates are indicated in each
plot. Input data from Suvarna et al. (2021) downloaded from the PRIDE Archive (PRIDE ID: PXD022296).
     
``` {r rocplot}
covid_prob_list <- test_models(
  model_list = covid_model_list,
  split_df = covid_split_df,
  type = "prob"
)
roc_plot(
  probability_list = covid_prob_list,
  split_df = covid_split_df,
  palette = "magma")
```
     
     
     
     
